Learning Heuristics of SAT by Graph Neural Networks

### 2018-12-15 report/paper.tex
report data

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.8]{beramono}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{xcolor,colortbl}
\usepackage{url}
\usepackage{listings}
\usepackage{paralist}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{flushend}
\usepackage{bcprules}
\usepackage{nicefrac}
\usepackage{xfrac}
\input{macros}

@@

We use PyTorch framework to implement the neural networks and reinforcement learning.
The code we use to implement message-passing neural network and gated-graph neural network 
are adapted from open-sourced repositories on Github 
\footnote{https://github.com/JamesChuanggg/ggnn.pytorch} 
\footnote{https://github.com/yujiali/ggnn}.
The policy network is a shallow neural network with three hidden layers.

@@

The training data and testing data are public and available online 
\footnote{https://www.cs.ubc.ca/~hoos/SATLIB/benchm.html}.
To measure the efficiency, we count the number of decisions made by baselines and the 
graph-based model.

@@

As artifical neural networks greatly succeeded in vision and natural language tasks in last ten years,
many works that combine deep learning and logical reasoning also has been emerged recently.

>>>>

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.8]{beramono}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{xcolor,colortbl}
\usepackage{url}
\usepackage{listings}
\usepackage{paralist}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{flushend}
\usepackage{bcprules}
\usepackage{nicefrac}
\usepackage{xfrac}
\usepackage{graphicx}
\input{macros}

@@

We use PyTorch framework to implement the neural networks and reinforcement learning.
The model we implemented is a gated-graph neural network, which is a variant of message-passing neural network.
The code we use to implement message-passing neural network and gated-graph neural network 
are adapted from open-sourced repositories on Github 
\footnote{https://github.com/JamesChuanggg/ggnn.pytorch} 
\footnote{https://github.com/yujiali/ggnn}.
The policy network is a shallow neural network with three hidden layers.

@@

As the first step, we train and test our models with 3-SAT problems consist of 
20 variables and 91 clauses. The training set contains 100 instances; we test
the models on 50 instances. For each training instance, we run 10 episodes and 
setup up a upper bound 400 steps for each episodes; but usually one episodes 
will finish in less than 100 steps.
The training data and testing data are public and available online 
\footnote{https://www.cs.ubc.ca/~hoos/SATLIB/benchm.html}.
To measure the efficiency, we count the number of decisions made by the DPLL algorithms 
with different baselines and the graph-based model. 
The intermediate states of DPLL such as unit propagation do not count as decisions.

Table 1 shows the basic statistical information (number of decisions) of 50 testing instances.
As we can see, compared with randomly picking a variable, 
the naive linear embedding already improves the performance.
However, we observe decline when using the graph neural network -- it has roughly the same
performance with the random stategy according the average stesp.

Figure 1 provides a detailed comparision between the three models based on the residual 
steps of these 50 test instances. In the graph noted in the form of "A vs. B", we obtain
the residual steps by computing $steps(B) - steps(A)$ for each instances.
For example, Figure 1 (a) shows the comparision between random strategy and RL with linear 
embedding, the negative numbers on y-axis mean that on these instances the RL model performs
better than the random strategy -- overall, RL is better on 52\% instances.
If we compare the random baseline with graph-based RL, again, we do not observe
significant improvement. We conjuecture this happens from two reasons: 
1) the graph model although captures the structural information, but it is more 
sophisticated, further parameter tunning (as well as for the policy network) 
may improve the performance;
2) If the model is trained with more data, we can expect a performance imporvement.

\begin{table}[]
\begin{tabular}{|l|l|l|l|}
\hline
           & mean  & std   & median \\ \hline
random strategy    & 24.66 & 22.56 & 21.5   \\ \hline
linear embedding RL & 19.94 & 12.90 & 15.0  \\ \hline
graph embedding RL  & 24.88 & 17.79 & 19.5  \\ \hline
\end{tabular}
\caption{Basic statistical information (steps)}
\end{table}

\begin{figure}[ht!]
  \centering
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \includegraphics[scale=0.3]{random_vs_linear.png}
    \caption{random strategy vs. linear embedding RL}
  \end{subfigure}

  \begin{subfigure}[b]{0.5\textwidth}
    \centering
  \includegraphics[scale=0.3]{random_vs_graph.png}
  \caption{random strategy vs. graph embedding RL}
  \end{subfigure}

  \caption{Detailed comparison of three models on 50 test instances.}
\end{figure}

@@

As artifical neural networks greatly succeeded in vision and natural language tasks in last ten years,
many works that combine deep learning and logical reasoning also has been emerged recently.
Some of the work are completely based on neural network, as opposed to utilizing
an existing deterministic framework. \citeauthor{selsam2018learning} proposed to 
use message-passing neural network to learn the assignment of a SAT problem \cite{selsam2018learning}.
We adapt a similar graph representation from \citeauthor{selsam2018learning}, but use it 
on DPLL, which poses a difficulty on handling dynamic graph strcuture.
Based on \citeauthor{selsam2018learning}'s NeuraSAT, a ICLR 2019 paper
further improves the model that directly generates the assignment in a unsupervised fashion, 
but works on circuit-sat problem\cite{anonymous2019learning}.
